### 2015-09-02 LaTeX/skewnormal.tex
few minor edits, almost ready to submit.

\newpage
\section{Introduction}\label{s:intro}
In many climatological applications, researchers are interested in learning about the average behavior of different climate variables (e.g. ozone, temperature, rainfall).
Our study is motivated by an air pollution application where the focus is not on the average behavior, but instead the behavior over a fixed threshold determined by government regulation.
More specifically, we consider the case of compliance for ozone.
A site is said to be in compliance if the fourth highest daily maximum 8-hour concentration averaged over 3 years does not exceed 75 parts per billion (ppb).
Figure \ref{fig:ozone-10jul} shows the ozone levels from July 10, 2005, at 1089 stations across the United States.
We see a large area above the compliance level in the midwest covering Ohio, Indiana, Illinois, and parts of the surrounding states.
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{plots/ozone-10jul-us.pdf}
  \caption{Ozone values on July 10, 2005}
  \label{fig:ozone-10jul}
\end{figure}

@@

Although any positive definite correlation function could be used, we choose to use the stationary isotropic \Matern correlation with
\begin{align}
  \text{cor}[v(\bs_1), v(\bs_2)] = \gamma I(\bs_1 = \bs_2) + (1 - \gamma)  \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right) \label{eq:matern}
\end{align}
where $\rho$ is the spatial range, $\nu$ is the smoothness, $\gamma$ is the proportion of variance accounted for by the spatial variation, $K_\nu$ is a modified Bessel function of the second kind, and $h = || \bs_1 - \bs_2 ||$.

@@

\subsection{Design}\label{s:simdesign}
For all simulation designs, we generated data from model (\ref{eq:partition})) in Section \ref{s:part} using $n_s=144$ sites and $n_t=50$ independent days.
The sites were generated Uniform$([0, 10] \times [0, 10])$.
We generated data from 4 different simulation designs:
\begin{enumerate} \setlength{\itemsep}{-0.5em}
  \item Gaussian marginal, $K=1$ knot
  % \item Symmetric-$t$ marginal, $K=1$ knot
  % \item Symmetric-$t$ marginal, $K=5$ knots
  \item Skew-$t$ marginal, $K=1$ knots
  \item Skew-$t$ marginal, $K=5$ knots
  \item Max-stable
  % \item Transformation below $T = q(0.80)$
\end{enumerate}
In the first three designs, the $v_t(\bs)$ terms were generated using a \Matern covariance with smoothness parameter $\nu = 0.5$ and spatial range $\rho = 1$.
For the covariance matrices in designs 1 -- 3, the proportion of the variance accounted for by the spatial variation was $\gamma = 0.9$ while the proportion of the variance accounted for by the nugget effect was $0.1$.
In the first design, $\sigma^2 = 2$ was used for all days which results in a Gaussian distribution.
For designs 2 and 3, $\sigma^2_{tk} \iid \text{IG}(3, 8)$ to give a $t$ distribution with 6 degrees of freedom.
For design 1, we set $\lambda = 0$.
For designs 2 and 3, $\lambda = 3$ was used as to simulate moderate skewness, and the $z_t$ were generated as described in (\ref{eq:sitezsig}).
In designs 1 -- 3, the mean $\bX^T \bbeta = 10$ was assumed to be constant across space.
In the fourth design, we generated from a spatial max-stable distribution \citep{Reich2012}.
In this design, data have marginal distributions that follow a generalized extreme value distribution with location parameter $1$, scale parameter $1$, and shape parameter $0.2$.
In this model, a random effect was used to induce spatial dependence using 144 spatial knots on a regular lattice in the square $[1, 9] \times [1, 9]$.
For this setting, we set $\gamma = 0.5$.
% In the final design, we generate $\tilde{y}$ using the setting from design 2, and then transform the data
% \begin{align}
%   y = \left\{ \begin{array}{lc}
%     \tilde{y}, \quad & \tilde{y} > T \\[0.5em]
%     T \exp\{\tilde{y} - T\}, \quad & \tilde{y} \le T
%   \end{array}\right.
% \end{align}
% where $T = q(0.80)$ is the 80th sample quantile of the data.
% The final design is included to explore the importance of using a threshold exceedance model when the distribution for the bulk of the data is misspecified.

@@

\section{Data analysis}\label{s:analysis}
We consider daily observations of maximum 8-hour ozone measurements for the 31 days of July 2005 at 1,089 Air Quality System (AQS) monitoring sites in the United States as the response (see Figure \ref{fig:ozone-10jul}).
For each site, we also have covariate information containing the estimated ozone from the Community Multi-scale Air Quality (CMAQ) modeling system.
Initially, we fit a linear regression assuming a mean function of
\begin{align}
  \text{E}[Y_i(\bs)] = \beta_0 + \beta_1 \cdot \text{CMAQ}_t(\bs). \label{eq:datamean}
\end{align}
Figure \ref{fig:ozone-qq} shows a Q-Q plot of the residuals compared to a skew-$t$ distribution with 10 d.f. and $\alpha = 1$.
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{plots/qq-res.pdf}
  \caption{Q-Q plot of the residuals for a skew-$t$ distribution with 10 d.f. and $\alpha = 1$ (right)}
  \label{fig:ozone-qq}
\end{figure}

@@

\appendix
\section{Appendices}
\subsection{MCMC details} \label{a:mcmc}
The MCMC sampling for the model \ref{s:hier} is done using {\tt R} (http://www.r-project.org). Whenever possible, we select conjugate priors (see Appendix \ref{a:posterior}); however, for some of the parameters, no conjugate prior distributions exist.
When no conjugate prior distribution exists, we use a random walk Metropolis-Hastings update step.
In each Metropolis-Hastings update, we tune the algorithm to give acceptance rates near 0.40.

@@

\subsection{Posterior distributions} \label{a:posterior}
\input{zpost.tex}
\input{betapost.tex}
\input{sigpost.tex}
\input{lambdapost.tex}

@@

\subsection{Skew-$t$ distribution} \label{a:skewt}
\subsubsection*{Univariate extended skew-$t$ distribution}
We say that $Y$ follow a univariate extended skew-$t$ distribution with location $\xi \in \calR$, scale $\omega > 0$, skew parameter $\alpha \in \calR$, extended parameter $\tau \in \calR$, and degrees of freedom $\nu$ if has distribution function
\begin{align}
  f_{\text{EST}}(y) = \omega^{-1} \frac{ f_T (z; \nu) }{ F_T(\tau / \sqrt{ 1 + \alpha^2}; \nu)} F_T\left[ (\alpha z + \tau) \sqrt{ \frac{ \nu + 1 }{ \nu + z^2}}; 0, 1, \nu + 1 \right]
\end{align}
where $f_T(t; \nu)$ is a univariate Student's $t$ with $\nu$ degrees of freedom, $F_T(t; \nu) = P(T < t)$, and \hbox{$z = (y - \xi) / \omega$}.
In the case that $\tau = 0$, then $Y$ follows a univariate skew-$t$ distribution.

@@

\subsubsection*{Proof that $\lim_{h \rightarrow \infty} \chi(h) > 0$}
Consider the bivariate distribution of $\bY = [Y(\bs), Y(\bt)]^T$, with $\varrho(h)$ given by (\ref{eq:matern}).
So, $\lim_{h \rightarrow \infty} \varrho(h) = 0$.
Then
\begin{align}
  \lim_{h \rightarrow \infty} \chi(h) = \bar{F}_{\text{EST}}\left\{ \sqrt{\eta + 1}; 0, 1, \alpha_1, \tau_1, \eta + 1 \right\} + \bar{F}_{\text{EST}}\left\{ \sqrt{\eta + 1}; 0, 1, \alpha_2, \tau_2, \eta + 1 \right\}.
\end{align}
Because the extended skew-$t$ distribution is not bounded above, for all $\bar{F}_{\text{EST}}(x) = 1 - F_{\text{EST}} > 0$ for all $x < \infty$.
Therefore, for a skew-$t$ distribution, $\lim_{h \rightarrow \infty} \chi(h) > 0$.

>>>>

\newpage
\section{Introduction}\label{s:intro}
% In many climatological applications, researchers are interested in learning about the average behavior of different climate variables (e.g. ozone, temperature, rainfall).
Epidemiological studies have linked air quality to public health concerns regarding morbidity and mortality \citep{Samet2000}.
As a results, the Environmental Protection Agency (EPA) has developed a set of standards to help reduce air pollution thereby improving air quality.
Our study is motivated by an air pollution application where the focus is not on the average behavior, but instead the behavior over a fixed threshold determined by government regulation.
More specifically, we consider the case of compliance for ozone.
A site is said to be in compliance if the fourth highest daily maximum 8-hour concentration averaged over 3 years does not exceed 75 parts per billion (ppb).
Figure \ref{fig:ozone-10jul} shows the ozone levels from July 10, 2005, at 1089 stations across the United States.
We see a large area above the compliance level in the midwest covering Ohio, Indiana, Illinois, and parts of the surrounding states.
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{plots/ozone-10jul-us.pdf}
  \caption{Ozone values on July 10, 2005}
  \label{fig:ozone-10jul}
\end{figure}

@@

Although any positive definite correlation function could be used, we choose to use the stationary isotropic \Matern correlation with
\begin{align}
  \text{cor}[v(\bs_1), v(\bs_2)] = \gamma I(\bs_1 = \bs_2) + (1 - \gamma)  \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right) \label{eq:matern}
\end{align}
where $\rho$ is the spatial range, $\nu$ is the smoothness, $\gamma$ is the proportion of variance accounted for by the spatial variation, $K_\nu$ is a modified Bessel function of the second kind, and $h = || \bs_1 - \bs_2 ||$.
We use this parameterization for spatial correlation because the $\gamma$ parameter permits the inclusion of a nugget effect to account for non-spatial variability due to issues like measurement error.

@@

\subsection{Design}\label{s:simdesign}
For all simulation designs, we generated data from model (\ref{eq:partition})) in Section \ref{s:part} using $n_s=144$ sites and $n_t=50$ independent days.
The sites were generated Uniform$([0, 10] \times [0, 10])$.
We generated data from 4 different simulation designs:
\begin{enumerate} \setlength{\itemsep}{-0.5em}
  \item Gaussian marginal, $K=1$ knot
  % \item Symmetric-$t$ marginal, $K=1$ knot
  % \item Symmetric-$t$ marginal, $K=5$ knots
  \item Skew-$t$ marginal, $K=1$ knots
  \item Skew-$t$ marginal, $K=5$ knots
  \item Max-stable
  % \item Transformation below $T = q(0.80)$
\end{enumerate}
In the first three designs, the $v_t(\bs)$ terms were generated using a \Matern covariance with smoothness parameter $\nu = 0.5$ and spatial range $\rho = 1$.
For the covariance matrices in designs 1 -- 3, the proportion of the variance accounted for by the spatial variation was $\gamma = 0.9$ while the proportion of the variance accounted for by the nugget effect was $0.1$.
In the first design, $\sigma^2 = 2$ was used for all days which results in a Gaussian distribution.
For designs 2 and 3, $\sigma^2_{tk} \iid \text{IG}(3, 8)$ to give a $t$ distribution with 6 degrees of freedom.
For design 1, we set $\lambda = 0$.
For designs 2 and 3, $\lambda = 3$ was used as to simulate moderate skewness, and the $z_t$ were generated as described in (\ref{eq:sitezsig}).
In designs 1 -- 3, the mean $\bX^T \bbeta = 10$ was assumed to be constant across space.
In the fourth design, we generated from a spatial max-stable distribution \citep{Reich2012}.
In this design, data have marginal distributions that follow a generalized extreme value distribution with location parameter $1$, scale parameter $1$, and shape parameter $0.2$.
In this model, a random effect was used to induce spatial dependence using 144 spatial knots on a regular lattice in the square $[1, 9] \times [1, 9]$.
For this setting, $\gamma_{\text{RS}} \in (0, 1)$, is similar to $\gamma$ in that it controls the relative strength of the nugget effect.
We set $\gamma_{\text{RS}} = 0.5$, which represents moderate spatial dependence.
% In the final design, we generate $\tilde{y}$ using the setting from design 2, and then transform the data
% \begin{align}
%   y = \left\{ \begin{array}{lc}
%     \tilde{y}, \quad & \tilde{y} > T \\[0.5em]
%     T \exp\{\tilde{y} - T\}, \quad & \tilde{y} \le T
%   \end{array}\right.
% \end{align}
% where $T = q(0.80)$ is the 80th sample quantile of the data.
% The final design is included to explore the importance of using a threshold exceedance model when the distribution for the bulk of the data is misspecified.

@@

\section{Data analysis}\label{s:analysis}
We consider daily observations of maximum 8-hour ozone measurements for the 31 days of July 2005 at 1,089 Air Quality System (AQS) monitoring sites in the United States as the response (see Figure \ref{fig:ozone-10jul}).
For each site, we also have covariate information containing the estimated ozone from the Community Multi-scale Air Quality (CMAQ) modeling system.
Initially, we fit a linear regression assuming a mean function of
\begin{align}
  \text{E}[Y_i(\bs)] = \beta_0 + \beta_1 \cdot \text{CMAQ}_t(\bs). \label{eq:datamean}
\end{align}
Figure \ref{fig:ozone-qq} shows a Q-Q plot of the residuals compared to a skew-$t$ distribution with 10 d.f. and $\alpha = 1$.
\begin{figure}
  \centering
  \includegraphics[width=0.5\linewidth]{plots/qq-res.pdf}
  \caption{Q-Q plot of the residuals for a skew-$t$ distribution with 10 d.f. and $\alpha = 1$ (right)}
  \label{fig:ozone-qq}
\end{figure}

@@

\appendix
\section{Appendices}
\subsection{MCMC details} \label{a:mcmc}
The MCMC sampling for the model \ref{s:hier} is done using {\tt R} (http://www.r-project.org). Whenever possible, we select conjugate priors (see Appendix \ref{a:posterior}); however, for some of the parameters, no conjugate prior distributions exist.
For these parameters, we use a random walk Metropolis-Hastings update step.
In each Metropolis-Hastings update, we tune the algorithm during the burn-in period to give acceptance rates near 0.40.

@@

\subsection{Posterior distributions} \label{a:posterior}

\subsubsection*{Conditional posterior of $z_{tk} \mid \ldots $}\label{s:mvcondu}
If knots are independent over days, then the conditional posterior distribution of $|z_{tk}|$ is conjugate.
For simplicity, drop the subscript $t$, let $\tilde{z}_{tk} = |z_{tk}|$, and define
\begin{align*}
R(\bs) = \left\{
    \begin{array}{ll}
        Y(\bs) - X(\bs) \beta &s \in P_l\\[1em]
        Y(\bs) - X(\bs) \beta - \lambda \tilde{z}(\bs) \qquad & s \notin P_l
    \end{array}
\right.
\end{align*}
Let
\begin{align*}
    R_1 &= \text{the vector of } R(\bs) \text{ for } s \in P_l \\
    R_2 &= \text{the vector of } R(\bs) \text{ for } s \notin P_l \\
    \Omega &= \Sigma^{-1}.
\end{align*}
Then
\begin{align*}
    \pi(z_l | \ldots) &\propto \exp \left\{ -\frac{ 1 }{ 2 } \left[
        \left( \begin{array}{c}
            R_1 - \lambda \tilde{z}_l \bOne\\
            R_2
        \end{array} \right)^T
        \left( \begin{array}{cc}
            \Omega_{11} & \Omega_{12}\\
            \Omega_{21} & \Omega_{22}
        \end{array} \right)
        \left( \begin{array}{c}
            R_1 - \lambda \tilde{z}_l \bOne\\
            R_2
        \end{array} \right)
        +  \frac{ {\tilde{z}_l}^2 }{ \sigma_l^2 }\right]
    \right\} I(z_l > 0) \\
        &\propto \exp \left\{ -\frac{ 1 }{ 2 } \left[ \Lambda_l {\tilde{z}_l}^2 - 2 \mu_l \tilde{z}_l \right] \right\}
\end{align*}
where
\begin{align*}
    \mu_l &= \lambda ( R_1^T \Omega_{11} + R_2^T \Omega_{21} )\bOne\\
    \Lambda_l &= \lambda^2 \bOne^T \Omega_{11} \bOne + \frac{ 1 }{ \sigma^2_l }.
\end{align*}
Then $\tilde{Z}_l | \ldots \sim N_{(0, \infty)} (\Lambda_l^{-1} \mu_l, \Lambda_l^{-1})$

\subsection*{Conditional posterior of $\beta \mid \ldots$}\label{s:betapost}
Let $\beta \sim \mbox{N}_{p}(0, \Lambda_0)$ where $\Lambda_0$ is a precision matrix. Then
\begin{align*}
    \pi(\beta \mid \ldots) & \propto \exp \left\{ - \frac{ 1 }{ 2 } \beta^T \Lambda_0 \beta - \frac{ 1 }{ 2 } \sum_{t = 1 }^T [\bY_t - X_t \beta - \lambda |z_t|]^T \Omega [\bY_t - X_t \beta - \lambda |z_t|] \right\}\\
     & \propto \exp \left\{ -\frac{ 1 }{ 2 } \left[ \beta^T \Lambda_\beta \beta  - 2 \sum_{ t = 1 }^T [ \beta^T X_t^T \Omega (\bY_t - \lambda |z_t| )] \right] \right\}\\
     & \propto \mbox{N} ( \Lambda_\beta^{-1} \mu_\beta , \Lambda_\beta^{-1})
\end{align*}
where
\begin{align*}
    \mu_\beta &= \sum_{t=1}^{T} \left[ X_t^T \Omega (\bY_t - \lambda |z_t|) \right]\\
    \Lambda_\beta &= \Lambda_0 + \sum_{ t = 1 }^{ T} X_t^T \Omega X_t.
\end{align*}

\subsection*{Conditional posterior of $\sigma^2 \mid \ldots$}\label{s:sigpost}
In the case where $L = 1$ and temporal dependence is negligible, then $\sigma^2$ has a conjugate posterior distribution.
Let $\sigma_t^2 \iid \mbox{IG}(\alpha_0, \beta_0)$. For simplicity, drop the subscript $t$. Then
\begin{align*}
    \pi(\sigma^2 \mid \ldots) & \propto (\sigma^2)^{ -\alpha_0 - 1 / 2 - n / 2 - 1} \exp \left\{ -\frac{\beta_0}{\sigma^2} - \frac{ |z|^2 }{2 \sigma^2} - \frac{ (\bY - \bmu)^T \Sigma^{-1} (\bY - \bmu) }{2 \sigma^2} \right\} \\
    & \propto (\sigma^2)^{ -\alpha_0 - 1 / 2 - n / 2 - 1} \exp \left\{ - \frac{ 1 }{ \sigma^2 } \left[\beta_0 + \frac{ |z|^2 }{ 2 } + \frac{ 1 }{ 2 } (\bY - \bmu)^T \Sigma^{-1} (\bY - \mu) \right] \right\} \\
    & \propto \mbox{IG} (\alpha^*, \beta^*)
\end{align*}
where
\begin{align*}
    \alpha^* &= \alpha_0 + \frac{1}{2} + \frac{n}{2} \\
    \beta^* &= \beta_0 + \frac{ |z|^2 }{ 2 } + \frac{ 1 }{ 2 }(\bY - \bmu)^T \Sigma^{-1} (\bY - \bmu).
\end{align*}
In the case that $L > 1$, a random walk Metropolis Hastings step will be used to update $\sigma^2_{lt}$.

\subsection*{Conditional posterior of $\lambda \mid \ldots$}\label{s:lambdapost}
For convergence purposes we model $\lambda = \lambda_1 \lambda_2$ where
\begin{align}
  \lambda_1 &= \left\{ \begin{array}{ll}
      +1 \quad & \text{w.p.} 0.5\\
      -1 \quad & \text{w.p.} 0.5
   \end{array}\right.\\
   \lambda^2_2 & \sim IG(\alpha_\lambda, \beta_\lambda).\\
\end{align}
Then
\begin{align*}
  \pi(\lambda_2 \mid \ldots) &\propto {\lambda_2^2}^{(-\alpha_\lambda - 1)} \exp\left\{ - \frac{ \beta_\lambda }{ \lambda_2^2} \right\} \prod_{t = 1}^{T} \prod_{k = 1}^K \frac{1} {\lambda_2} \exp\left\{ - \frac{ z^2_{tk} }{ 2 \lambda_2^2  \sigma_{tk})^2} \right\} \\
   & \propto {\lambda_2^2}^{(-\alpha_\lambda - kt - 1)} \exp \left\{ -\frac{1}{\lambda^2_2} \left[ \beta_\lambda + \frac{ z^2 }{2 \sigma^2_{tk}} \right] \right\}
\end{align*}
Then $\lambda_2 \mid \ldots \sim IG \left(\alpha_\lambda + kt, \beta_\lambda + \frac{ z^2 }{2 \sigma^2_{tk}} \right)$

@@

\subsection{Skew-$t$ distribution} \label{a:skewt}
\subsubsection*{Univariate skew-$t$ distribution}
We say that $Y$ follows a univariate extended skew-$t$ distribution with location $\xi \in \calR$, scale $\omega > 0$, skew parameter $\alpha \in \calR$, and degrees of freedom $\nu$ if has distribution function
\begin{align}
  f_{\text{EST}}(y) = 2 f_T (z; \nu) F_T\left[ \alpha z \sqrt{ \frac{ \nu + 1 }{ \nu + z^2}}; \nu + 1 \right]
\end{align}
where $f_T(t; \nu)$ is a univariate Student's $t$ with $\nu$ degrees of freedom, $F_T(t; \nu) = P(T < t)$, and \hbox{$z = (y - \xi) / \omega$}.

@@

\subsubsection*{Proof that $\lim_{h \rightarrow \infty} \chi(h) > 0$}
Consider the bivariate distribution of $\bY = [Y(\bs), Y(\bt)]^T$, with $\varrho(h)$ given by (\ref{eq:matern}).
So, $\lim_{h \rightarrow \infty} \varrho(h) = 0$.
Then
\begin{align}
  \lim_{h \rightarrow \infty} \chi(h) = \bar{F}_{\text{EST}}\left\{ \sqrt{\eta + 1}; 0, 1, \alpha_1, \tau_1, \eta + 1 \right\} + \bar{F}_{\text{EST}}\left\{ \sqrt{\eta + 1}; 0, 1, \alpha_2, \tau_2, \eta + 1 \right\}.
\end{align}
Because the extended skew-$t$ distribution is not bounded above, for all $\bar{F}_{\text{EST}}(x) = 1 - F_{\text{EST} (x)} > 0$ for all $x < \infty$.
Therefore, for a skew-$t$ distribution, $\lim_{h \rightarrow \infty} \chi(h) > 0$.
